3D MAGNETIC MODELLING FOR ELIPSOIDS

### 2016-11-30 manuscript/Paper-magnetic-elipsoid_metodologia.tex
correcao no meu nome
Atualizacao introducao e metodologia

\Author[1]{Diego}{Takahashi Tomazella}
\Author[1]{Vanderlei}{Coelho Oliveira Junior}

@@

In the 19th century, \cite{dirichlet1839nouvelle} manage to solve the problem of the potential of a homogeneous ellipsoid with semi-axes $a \ge b \ge c$ by using the first and second Legendre's normal elliptic integrals. Later, \cite{maxwell1881treatise} showed that only bodies of second degree can be uniformly magnetized when placed in a uniform field. He also presented the formulae of the demagnetizing field in a uniformly magnetized ellipsoid of revolution.

The formulas for the demagnetizing factors of the general ellipsoid alongside a table of values for different proportions of semi-axis was published by \cite{osborn1945demagnetizing}. In the same year, \cite{stoner1945xcvii} presented a more simple and friendly formulas of the demagnetizing factors for an ellipsoid, with step-by-step deduction. This work includes formulas for different parametrizations for the ellipsoid (triaxial, prolate, oblate).

@@

In the decade of 1970 \cite{farrar1979some} showed in his work the value of the ellipsoidal model in geophysic exploration, where it was used for analysis of magnetic anomalies in Tennant Creeks's gold mine in Australia. The ellipsoid was the most correct geometry to model bodies of lenticular pipe shape and the most appropriate to handle self-desmagnetization of bodies of high susceptibilty - magnetic susceptibility higher than 0.1 SI \citep{clark2014methods}.
\cite{guo1998self}, showed three case studies of the importance of self-demagnetization of high susceptibilties bodies when modelling magnetic sources. A magnetic-iron deposit, volcanic-hosted iron deposit and an ultramafic-hoste nickel andcopper deposits.
The same author years later,\cite{guo2001systematic}, published another paper analysing the errors of what an uncorrected by demagnetization data can affect on anmalies interpretation.

The magnetic field resultant of an ellipsoid was presented the first time in full modeling in the space domain by \cite{emerson1985magnetic}, along a compilation of several others geometric shapes. Through a consistent notation this work tapped a hole in the literature that missed the ellipsoidal model. This model, however, consists in a division between prolate and oblate ellipsoids, with a few differences in the algorithm.
A generalized model for triaxials ellipsoids was published by \cite{clark1986magnetic}. Both in Emerson et al. and Clark et al. works, they were the firsts to include the calculations of the magnetic field in a different coordinate system other than with the position $(0,0,0)$ in the center of the ellipsoid (through a matrix of rotation). However, the codes presented for HP-41C are outdated and doesn't have a descrition of the algorithm, difficulting the reproduction by the cientific commnunity. This may the reason why the magnetic field modelling of the ellipsoid, in general, is only used through comercial softwares.

\section{Methodology}
The total-field anomaly can be described as a difference between the total field vector and the induced magnetization of the crust of the Earth, that includes any anomalous distribuition of magnetization that can occur by magnetizied bodies in subsurface. This difference can be written as:
\begin{equation} \label{eq:1}
\Delta T^0_i = ||T_i|| - ||F_i||,
\end{equation}
where $\Delta T^0_i$ is the observed vector of total-field anomaly in the $i$-th position ($x_i$, $y_i$, $z_i$), $i = 1,...N$, $F_i$ is the geomagnetic field and $||.||$ is the Euclidian norm. The total-field vector $T_i$ is:
\begin{equation}\label{eq:2}
T_i = F_i + B_i,
\end{equation}
and $B_i$ is the total induced magnetization vector produced by all sources of anomalous susceptibility distribuition.
Considering $F_i$ a constant vector, $F_0$ for local ou regional scale, and that $||F_0|| \gg ||B_0||$, since, $B_i$ is a small pertubation of the magnetic field $F_i$, we can approximate the euclidian norm of the vector $T_i$ by a Taylor's expansion of first order:
\begin{equation}\label{eq:3}
||T_i|| \approx ||F_0 + B_i|| \approx ||F_0|| + F^T B_i
\end{equation}
where $T$ indicates transposition and
\begin{equation}\label{eq:4}
\hat{F} = \dfrac{F_o}{||F_0||}
\end{equation}

is a unit vector that represents the gradient of the function $||T_i||$ in relation to the vector's components $T_i$. This way, we can approximate the Eq.(\ref{eq:1}) of total-field anomaly to:
\begin{equation}\label{eq:5}
\Delta T \approx \hat{F^T} B_i, \quad i=1,...,N
\end{equation}

@@

This new coordinate system ($x_1,x_2,x_3$) is defined by the unit vectors $\hat{v}_h (h=1,2,3)$ with the respect with the geographic axis $x$, $y$ e $z$:
\begin{equation}
\hat{v_1} = (l_1,m_1,n_1) = (-\cos\alpha \cos\delta, -\sin\alpha \cos\delta, -\sin\delta)
\end{equation}

\begin{equation}
\begin{split}
\hat{v_2} = (l_2,m_2,n_2) = (\cos\alpha \cos\gamma \sin\delta + \sin\alpha \sin\gamma,\\ \sin\alpha \cos\gamma \sin\delta - \cos\alpha \sin\gamma, -\cos\gamma \cos\delta)
\end{split}
\end{equation}

\begin{equation}
\begin{split}
\hat{v_3} = (l_3,m_3,n_3) = (\sin\alpha \cos\gamma - \cos\alpha \sin\gamma \sin\delta,\\ -\cos\alpha \cos\gamma - \sin\alpha \sin\gamma \sin\delta, \sin\gamma \cos\delta)
\end{split}
\end{equation}

@@

\begin{equation}
p_0 = a^2b^2c^2 - b^2c^2x_1^2 - c^2a^2x_2^2 - a^2b^2x_3^2
\end{equation}

@@

\begin{equation}
\tilde{J}_R = K\tilde{F} + \tilde{J}_{NRM}
\end{equation}

@@

\begin{equation}
\tilde{J}_{Rc} = A^{-1} \tilde{J}_R
\end{equation}

>>>>

\Author[1]{Diego}{Takahashi Tomazella}
\Author[1]{Vanderlei}{C. Oliveira Jr}

@@

In the 19th century, \cite{dirichlet1839nouvelle} manage to solve the problem of the potential of a homogeneous ellipsoid with semi-axes $a \ge b \ge c$ by using the first and second Legendre's normal elliptic integrals. Later, \cite{maxwell1881treatise} showed that only bodies of second degree can be uniformly magnetized when placed in a uniform field. \cite{thomson1879and} presented the formulas of the potential and demagnetizing field using the standard integrals expressions for the ellispoid of revolution.

The formulas for the demagnetizing factors of the general ellipsoid alongside a table of values for different proportions of semi-axis was published by \cite{osborn1945demagnetizing}. In the same year, \cite{stoner1945xcvii} presented a more simple and friendly formulas of the demagnetizing factors for an ellipsoid, with step-by-step deduction. This work includes formulas for different parametrizations of the ellipsoid (triaxial, prolate, oblate).

@@

In the decade of 1970, \cite{farrar1979some} showed in his work the value of the ellipsoidal model in geophysic exploration, where it was used for analysis of magnetic anomalies in Tennant Creeks's gold mine in Australia. The ellipsoid was the most correct geometry to model bodies of lenticular pipe shape and the most appropriate to handle self-desmagnetization of bodies of high susceptibilty - magnetic susceptibility higher than 0.1 SI \citep{clark2014methods}.
\cite{guo1998self}, showed three case studies of the importance of self-demagnetization of high susceptibilties bodies when modelling magnetic sources. A magnetic-iron deposit, a volcanic-hosted iron deposit and an ultramafic-hosted nickel and copper deposits.
The same author years later,\cite{guo2001systematic}, published another paper analysing the errors of what an uncorrected by demagnetization data can affect in anomalies interpretation.

The magnetic field resultant of an ellipsoid was presented the first time in full modeling in the space domain by \cite{emerson1985magnetic}, along a compilation of several others geometric shapes. Through a consistent notation this work tapped a hole in the literature that missed the ellipsoidal model. This model, however, consists in a division between prolate and oblate ellipsoids, with a few differences in the algorithm.
A generalized model for triaxials ellipsoids was published by \cite{clark1986magnetic}. Both in Emerson et al. and Clark et al. works, they were the firsts to include the calculations of the magnetic field in a different coordinate system other than with the origin in the center of the ellipsoid (through a matrix of rotation). However, the codes presented for HP-41C calculators are outdated and don't have a descrition of the algorithm, difficulting the reproduction by the cientific commnunity. This may be the reason why the magnetic field modelling of the ellipsoid, in general, is manly used through commercial softwares.

As shown, the problem of the magnetic field of the ellipsoid is scattered through the literature, wich can difficult the access to many. This article come as a junction of previous works, making the problem easier and more understandable. Together we present an algorithm in a modern language with a complete descrition of it. The code will be available for free usage of the geophysics community.

\section{Methodology}

The total-field anomaly can be described as a difference between the total field vector and the induced magnetization of the crust of the Earth, that includes any anomalous distribuition of magnetization that can occur by magnetizied bodies in subsurface. This difference can be written as:
\begin{equation} \label{eq:1}
\Delta T^0_i = ||T_i|| - ||F_i||,
\end{equation}
where $\Delta T^0_i$ is the observed vector of total-field anomaly in the $i$-th position ($x_i$, $y_i$, $z_i$), $i = 1,...N$, $F_i$ is the geomagnetic field and $||.||$ is the Euclidian norm. The total-field vector $T_i$ is:
\begin{equation}\label{eq:2}
T_i = F_i + B_i,
\end{equation}
and $B_i$ is the total induced magnetization vector produced by all sources of anomalous susceptibility distribuition.
Considering $F_i$ a constant vector, $F_0$ for local ou regional scale, and that $||F_0|| \gg ||B_0||$, since, $B_i$ is a small pertubation of the magnetic field $F_i$, we can approximate the euclidian norm of the vector $T_i$ by a Taylor's expansion of first order:
\begin{equation}\label{eq:3}
||T_i|| \approx ||F_0 + B_i|| \approx ||F_0|| + F^T B_i
\end{equation}
where $T$ indicates transposition and
\begin{equation}\label{eq:4}
\hat{F} = \dfrac{F_o}{||F_0||}
\end{equation}

is a unit vector that represents the gradient of the function $||T_i||$ in relation to the vector's components $T_i$. This way, we can approximate the Eq.(\ref{eq:1}) of total-field anomaly to:
\begin{equation}\label{eq:5}
\Delta T \approx \hat{F}^T \bar{B}_i, \quad i=1,...,NF
\end{equation}

The field now $\bar{B}_i$ is represented this way to show that is in the geographic coordinates. One of the main steps to calculate the magnetic field of an ellipsoid is to change this system of coordinates to the body coordinates with origin in the center of it. The conversion can be done through a matrix of rotation $\bar{\bar{V}}$ that differs with the geometric type of ellipsoid, so the formulas will be shown later, separately for each one. But, in a general way, the conversion is:
\begin{equation}\label{eq:5}
\bar{r}_i = \bar{\bar{V}} \tilde{r}_i + \bar{r}_c,
\end{equation}

The vector $\bar{r}_i$ has the positions $x,y,z$ in the geographic coordinates, $\bar{r}_i$ has the positions $x1,x2,x3$ in the body coordinates and $\bar{r}_c$ is the position of the center of the body in the geographic coordinates.

The same way the matrix of conversion can be used to change the magnetic field calculate in the body coordinates, back to the geopraphic, where is more useful to us:
\begin{equation}\label{eq:5}
\bar{B}_i = \bar{\bar{V}} \tilde{B}_i
\end{equation}

@@

This new coordinate system ($x_1,x_2,x_3$) is defined by the unit vectors $\hat{v}_h (h=1,2,3)$ with the respect with the geographic axis $x$, $y$ e $z$:
\begin{equation}
\textbf{v}_1 = (l_1,m_1,n_1) = (-\cos\alpha \cos\delta, -\sin\alpha \cos\delta, -\sin\delta)
\end{equation}

\begin{equation}
\begin{split}
\textbf{v}_2 = (l_2,m_2,n_2) = (\cos\alpha \cos\gamma \sin\delta + \sin\alpha \sin\gamma,\\ \sin\alpha \cos\gamma \sin\delta - \cos\alpha \sin\gamma, -\cos\gamma \cos\delta)
\end{split}
\end{equation}

\begin{equation}
\begin{split}
\textbf{v}_3 = (l_3,m_3,n_3) = (\sin\alpha \cos\gamma - \cos\alpha \sin\gamma \sin\delta,\\ -\cos\alpha \cos\gamma - \sin\alpha \sin\gamma \sin\delta, \sin\gamma \cos\delta)
\end{split}
\end{equation}

@@

Where:
\begin{equation}
p_0 = a^2b^2c^2 - b^2c^2x_1^2 - c^2a^2x_2^2 - a^2b^2x_3^2
\end{equation}

@@

\begin{equation}
\textbf{J}_R = K\textbf{F} + \textbf{J}_{NRM}
\end{equation}

@@

\begin{equation}
\textbf{J}_{Rc} = A^{-1} \textbf{J}_R
\end{equation}
